Compilation With Program Synthesis

### 2023-08-22 new-paper/introduction.tex
bibliography; a few sentences on layout polymorphism

Synthetic separation logic (SSL) is the specification language for the \SuSLik{} program synthesizer. \SuSLik{} will take a precondition and postcondition pair together
with a function prototype and use these to synthesize a function that satisfies that specification in a C-like language called \SuSLang.

@@

>>>>

Synthetic separation logic (SSL) is the specification language for the \SuSLik{} program synthesiser. \SuSLik{} will take a precondition and postcondition pair together
with a function prototype and use these to synthesise a function that satisfies that specification in a C-like language called \SuSLang.

@@

Two versions of \Pika{} have been built. \Pika{} 1 has a focus on using \SuSLik{} to generate all of the low-level code, while \Pika{} 2 adds an additional synthesis feature inside the \Pika{} language itself and has a builtin testing framework. In the present work, we will start by talking about \Pika{} 1, describe the fundamental limitations of using \SuSLik{} to generate everything and then use this to motivate the creation of \Pika{} 2. The two versions are then compared.

### 2023-08-22 new-paper/limitations.tex
Bring more things over
bibliography; a few sentences on layout polymorphism

There is a significant limitation to our approach so far: \SuSLik will often fail to synthesize code from valid specifications. One instance of this problem is the following: In a certain class of functions, \SuSLik{} requires a different number of pointer indirections than it will with other functions. We have not found a way to perfectly characterize such functions and, since this will change the way the functions will need to be called, this represents a major impediment to our approach. It is worth noting that this is also not the \textit{only} situation in which \SuSLik{} fails to synthesize a program.

@@

>>>>

There is a significant limitation to our approach so far: \SuSLik{} will often fail to synthesize code from valid specifications. One instance of this problem is the following: In a certain class of functions, \SuSLik{} requires a different number of pointer indirections than it will with other functions. We have not found a way to perfectly characterize such functions and, since this will change the way the functions will need to be called, this represents a major impediment to our approach. It is worth noting that this is also not the \textit{only} situation in which \SuSLik{} fails to synthesize a program.

@@

Furthermore, \Pika{} 1 relies on a kind of \textit{implicit} layout polymorphism: The layouts used are all inferred. There are no type variables that represent layouts to constrain a type signature. Towards solving this, \Pika{} 2 also uses explicit layout passing. We have not yet implemented full layout polymorphism, but this can be built on top of explicit layout passing.

### 2023-08-22 new-paper/semantics.tex
bibliography; a few sentences on layout polymorphism

\subsection{Formal Semantics of \tool}
\label{sec:formal}

@@

\subsubsection{Overview of the Two Interpretations}

@@

% \subsubsection{Abstract Machine}
First, consider the abstract machine semantics. In this case, we are able to choose which branch of a layout to use by evaluating the expression we are applying it
          to until the expression is reduced to a constructor value (where a ``constructor value'' is either a value or a constructor applied to constructor values). If the expression is well-typed, this will always be a constructor
          of the algebraic data type corresponding to the layout. The two rules that this applies to are {\sc AM-Lower} and
          {\sc AM-Instantiate}.
        To interpret a function application, we interpret its arguments and substitute the results into the body
          of the function. We then proceed to interpret the substituted function body. This process is performed by the {\sc AM-Instantiate} rule.
  % }
% \\

% \subsubsection{SSL Translation}
Next, consider the SSL translation. Here we can determine which layout branch to use by generating a Boolean condition that will be
          true if and only if the SSL proposition on the right-hand side of the branch holds for the heap. Note
          that we assume that the programmer-supplied layout definitions are \textit{injective} functions from
          algebraic data types to SSL assertions (up to bi-implication).
We can directly translate function applications into SSL inductive predicate applications. Since inductive predicates
          already allow for recursive applications, there is no special handling necessary for recursion.
      % \end{enumerate}
  % }
% \\

@@

\subsubsection{Abstract Machine Semantics}
In this section, we will define an abstract machine semantics for \tool and relate this to
the standard semantics for SSL.

@@

\begin{figure}[b]
%\vspace{-5pt}
% \setlength{\belowcaptionskip}{-12pt}
% \setlength{\abovecaptionskip}{0pt}
\centering
% \begin{subfigure}{0.49\textwidth}
\begin{minipage}{1.0\linewidth}
{\footnotesize{
\[
  \begin{array}{c}
    \labinfer[C-Int]{e \concrete{\Int}}
      {\Gamma \vdash e : \Int}
    ~~~
    \labinfer[C-Bool]{e \concrete{\Bool}}
      {\Gamma \vdash e : \Bool}
    \\\\
    \labinfer[C-Layout]{e \concrete{\alpha}}
      {(A : \alpha \monic \layout[x]) \in \Sigma &
       \Gamma \vdash e : A
      }
  \end{array}
\]
}}
\end{minipage}
  \caption{Concreteness judgment rules}
  \label{fig:concreteness-rules}
% \end{subfigure}
%%
% \begin{subfigure}{0.49\textwidth}
\begin{minipage}{1.0\linewidth}
{\footnotesize{
\[
\begin{array}{c}
  \labinfer[G-Fn]{(f\; (C\; b_1 \cdots b_n) := e) \Rightarrow f : \beta \rightarrow \gamma}
    {(C : \alpha_1 \rightarrow \alpha_2 \rightarrow \cdots \rightarrow \alpha_n \rightarrow \beta) \in \Delta\\
    b_1 : \alpha_1, b_2 : \alpha_2, \cdots, b_n : \alpha_n \vdash e : \gamma
    }
\end{array}
\]
}}
\end{minipage}
  \caption{Global definition typing}
  \label{fig:globals-typing-rules}
% \end{subfigure}
%\setlength{\belowcaptionskip}{-15pt}
\caption{Rules for concreteness judgement and typing global definitions}
\label{fig:meh}
\end{figure}

>>>>

% \subsection{Formal Semantics of \tool}
% \label{sec:formal}

@@

\subsection{Overview of the Two Interpretations}

@@

% \subsubsection{Abstract Machine}
\paragraph{Abstract machine}
First, consider the abstract machine semantics. In this case, we are able to choose which branch of a layout to use by evaluating the expression we are applying it
          to until the expression is reduced to a constructor value (where a ``constructor value'' is either a value or a constructor applied to constructor values). If the expression is well-typed, this will always be a constructor
          of the algebraic data type corresponding to the layout. The two rules that this applies to are {\sc AM-Lower} and
          {\sc AM-Instantiate}.
        To interpret a function application, we interpret its arguments and substitute the results into the body
          of the function. We then proceed to interpret the substituted function body. This process is performed by the {\sc AM-Instantiate} rule.
  % }
% \\

% \subsubsection{SSL Translation}
\paragraph{SSL translation}
Next, consider the SSL translation. Here we can determine which layout branch to use by generating a Boolean condition that will be
          true if and only if the SSL proposition on the right-hand side of the branch holds for the heap. Note
          that we assume that the programmer-supplied layout definitions are \textit{injective} functions from
          algebraic data types to SSL assertions (up to bi-implication).
We can directly translate function applications into SSL inductive predicate applications. Since inductive predicates
          already allow for recursive applications, there is no special handling necessary for recursion.
      % \end{enumerate}
  % }
% \\

@@

\subsection{Abstract Machine Semantics}
In this section, we will define an abstract machine semantics for \tool and relate this to
the standard semantics for SSL.

@@

\begin{figure}[b]
%\vspace{-5pt}
% \setlength{\belowcaptionskip}{-12pt}
% \setlength{\abovecaptionskip}{0pt}
\centering
% \begin{subfigure}{0.49\textwidth}
\begin{minipage}{1.0\linewidth}
{\footnotesize{
\[
  \begin{array}{c}
    \labinfer[C-Int]{e \concrete{\Int}}
      {\Gamma \vdash e : \Int}
    ~~~
    \labinfer[C-Bool]{e \concrete{\Bool}}
      {\Gamma \vdash e : \Bool}
    \\\\
    \labinfer[C-Layout]{e \concrete{\alpha}}
      {(A : \alpha \monic \layout[x]) \in \Sigma &
       \Gamma \vdash e : A
      }
  \end{array}
\]
}}
\end{minipage}
  \caption{Concreteness judgment rules}
  \label{fig:concreteness-rules}
% \end{subfigure}
%%
% \begin{subfigure}{0.49\textwidth}
\begin{minipage}{1.0\linewidth}
{\footnotesize{
\[
\begin{array}{c}
  \labinfer[G-Fn]{(f\; (C\; b_1 \cdots b_n) := e) \Rightarrow f : \beta \rightarrow \gamma}
    {(C : \alpha_1 \rightarrow \alpha_2 \rightarrow \cdots \rightarrow \alpha_n \rightarrow \beta) \in \Delta\\
    b_1 : \alpha_1, b_2 : \alpha_2, \cdots, b_n : \alpha_n \vdash e : \gamma
    }
\end{array}
\]
}}
\end{minipage}
  \caption{Global definition typing}
  \label{fig:globals-typing-rules}
% % \end{subfigure}
% %\setlength{\belowcaptionskip}{-15pt}
% \caption{Rules for concreteness judgement and typing global definitions}
% \label{fig:meh}
\end{figure}
